---
title: "calm_fits"
author: "Victor Navarro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{calm_fits}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(calm)
library(ggplot2)
library(data.table)
theme_set(theme_bw())
data(pati)
set.seed(2022)
```

# Fitting HeiDI to empirical data

In this demo, I fit HeiDI to some empirical data (Patitucci et al., 2016, Experiment 1). 
This will involve writing a function that produces model responses organized as the empirical data, 
and using that function for maximum likelihood estimation (MLE). We begin with a short 
overview of the data, then move to the model function, and finally fit the model.

## The data

The data (`pati`) contains the responses (lever presses or lp, and nose pokes or np) 
for 32 subjects (rats) across 6 blocks of training (2 sessions per block). The animals 
were trained to associate each of two levers to one of two 
unconditioned stimuli (pellets or sucrose). Let's take a look at it.

```{r}
summary(pati)
pati |> ggplot(aes(x = block, y = rpert, colour = us)) +
  geom_line(aes(group = interaction(us, subject)), alpha = .3) +
  stat_summary(geom = "line", fun = "mean", linewidth = 1) +
  labs(x = "Block", y = "Responses per trial", colour = "US") +
  facet_grid(~response)
```

The thicker lines are group averages; the rest are individual subjects. 
We ignore the specific mapping between levers and USs here, because that was 
counterbalanced across subjects. However, the counterbalancing will end up being 
relevant (see ahead).

## Writing the model function

The biggest hurdle in fitting the model to empirical data is to write a function 
that, given a vector of parameters and an experiment generates responses that 
are organized as the empirical data. Let's begin by summarizing the data first, 
so we know what to aim for.

```{r}
pati_summ <- setDT(pati)[,
  list("rpert" = mean(rpert)),
  by = "block,us,response"
]
# set order (relevant for the future)
setorder(pati_summ, block, response, us)
head(pati_summ)
```

We now will prepare the experiment 
as you would pass to `run_experiment`experiment.

This is not a trivial issue because HeiDI, like many models, 
is sensitive to order effects. Hence, the arguments we prepare here must
reflect the behavior of the model after a "general" experimental 
procedure, and not the quirks of an unfortunate run of trials. Here, 
we simply address this issue by running several iterations of the experiment 
(with random trial orders) and averaging all experiments before evaluating the 
likelihood of the parameters.

So what do we have to design? The experiment presented by Patittuci 
et al. (2016) was fairly simple, and it can be reduced to the 
presentations of two levers, each followed by a different appetitive 
outcome. Here, we will assume that the two outcomes are independent 
from each other. We will also take some liberties with the number of trials 
we specify to reduce computing time.


```{r}
# The design data.frame
des_df <- data.frame(
  group = c("CB1", "CB2"),
  training = c(
    "12L>(Pellet)/12R>(Sucrose)",
    "12L>(Sucrose)/12R>(Pellet)"
  ),
  rand_train = FALSE
)
# The parameters
# the actual parameter values don't matter,
# as our function will re-write them inside the optimizer call
parameters <- get_parameters(des_df,
  model = "HD2022"
)
# The arguments
experiment <- make_experiment(des_df,
  parameters = parameters, model = "HD2022",
  iterations = 4
)
experiment
```

Note we specified two counterbalancings as groups. 
We must reproduce the counterbalancings in 
the data we are trying to fit as close as possible. Otherwise, 
the optimization process might latch onto experimentally-irrelevant variables. 
For example, it can be seen in `pati` that there was more lever pressing whenever 
a lever was paired with pellets. If we didn't counterbalance the identities of the 
levers and USs, the optimization might result in one of the levers being less 
salient than the other.

We can now begin to write the model function. First, it would be a good 
to see what results `run_experiment` returns.

```{r}
exp_res <- run_experiment(experiment)
results(exp_res)
```

Although `results` returns many model outputs, 
we only care about one of them: `rs` (the model responses).
With them, we can write our model function.

```{r}
my_model_function <- function(pars, exper) {
  # extract the parameters from the model
  new_parameters <- parameters(exper)[[1]]
  # assign alphas
  new_parameters$alphas[] <- pars
  # reassign parameters to the experiment
  parameters(exper) <- new_parameters # note parameters method
  # running the model and selecting rs
  exp_res <- run_experiment(exper)
  # summarizing the model
  rs <- results(exp_res)$rs
  # calculate extra variables
  rs$response <- ifelse(rs$s1 %in% c("Pellet", "Sucrose"), "np", "lp")
  rs$block <- ceiling(rs$trial / 4)
  # filtering
  rs <- rs[s2 %in% c("Pellet", "Sucrose") &
    (response == "np" | (response == "lp" &
      mapply(grepl, s1, trial_type)))]
  rs <- rs[, list(value = mean(value)), by = "block,s2,response"]
  rs$value
}
```

Let's dissect the function above in its three parts. 

1. We put `parameters` (the optimizer parameters) into the 
experiment using the `parameters` method.

2. We run the model and select the relevant information (rs).

3. We summarise the model responses and return them.[^1]

Let's see the function in action.

```{r}
my_model_function(c(.1, .2, .4, .3), experiment)
```

The order of the empirical data and model responses must match. 
I cannot emphasize this point enough: there is nothing within the fit function that checks 
or reorders the data for you. You are the sole responsible for making sure both of these 
pieces of data are in the same order. A simple way would be to print the model results 
before the return (see above). Once we have made sure everything is looking good, 
we can fit the model.

## Fitting the model

We fit models using the `fit_model` function. This function requires 4 arguments: 

1. The (empirical) data
2. A model function
3. The arguments with which to run the model function.
4. The optimizer options.

We have done a great job taking care of the first three, so let's tackle the last.

```{r}
my_optimizer_opts <- get_optimizer_opts(
  model_pars = names(parameters$alphas),
  optimizer = "ga",
  ll = c(0, 0, 0, 0),
  ul = c(1, 1, 1, 1),
  family = "normal"
)
my_optimizer_opts
```

The `get_optimizer_opts` function returns many things:

1. model_pars: The name of the model parameters. Here, the name of the alphas for each stimuli.
2. ll and ul: The lower and upper limits for the parameter search. Consider shrinking these to speed up the process.
3. optimizer: The numerical optimization technique we wish to use during MLE estimation.
4. family: The family distribution we assume for our model. In practice, what you request 
here will be used to determine the link function to transform model responses, 
and the likelihood function used in the objective function. The normal family here 
does nothing fancy to the model responses, and will estimate an extra parameter, 
scale, which scales the model responses into the scale of the empirical data. 
When it comes to likelihood functions, this family will use the normal density of 
the data and model differences.
5. family_pars: The family-specific parameter being estimated alongside salience parameters.
6. verbose: Whether to print parameters and objective function values as we optimize.

You are free to modify these; just make sure the structure of the list returned 
by `get_optimizer_opts` remains the same.

We can also pass extra parameters to the optimizer call we 
are using (e.g., the `par` argument for `optim`, or `parallel` for `ga`).
Here, we fit the model in parallel with `ga`, and for only 10 iterations.

And with that, we can fit the model! (be patient if you are following along)

```{r, eval=FALSE, include=TRUE}
the_fit <- fit_model(pati_summ$rpert,
  model_function = my_model_function,
  exper = experiment,
  optimizer_options = my_optimizer_opts,
  maxiter = 10,
  parallel = TRUE
)
```

```{r, include = F, echo = F}
# save("the_fit", file = "vignettes/calm_fits_fit.rda")
# load(file = "vignettes/calm_fits_fit.rda")
load(file = "calm_fits_fit.rda")
```


The `fit_model` function returns a lot of information to track what we put in and what we got out. 
However, typing the model in the console will show the MLE parameters we obtained this time 
and their negative log likelihood, given the data:

```{r}
the_fit
```

That's good and all, but how well does a model run with those parameters "visually" fit the data? We can obtain the predictions from the model via the `predict` function.

```{r}
pati_summ$prediction <- predict(the_fit, exper = experiment)
pati_summ[, data := rpert][, rpert := NULL]
pati_summ <- melt(pati_summ, measure.vars = c("prediction", "data"))
pati_summ |>
  ggplot(ggplot2::aes(
    x = block, y = value,
    colour = us,
    linetype = variable
  )) +
  geom_line() +
  theme_bw() +
  facet_grid(us ~ response)
```

This looks pretty good! Save from some blatant misfits, of course. Now you know 
everything you need to fit calm to your empirical data. Go forth!

### A final note
This vignette was pre-generated, as I don't want the user to fit the 
model at the time of installation. I will try to keep up with it as the 
package develops, but if you spot any inconsistencies, please drop me a line.

[^1]: Within this step, we also filter all output nodes that are not related to expecting one of the USs, 
we classify responses as being nosepokes (produced by the US) or 
lever presses (produced by the levers), and calculate the mean across blocks of trials. 
Fitting a model is no trivial task!
