---
title: "calmr_basics"
author: "Victor Navarro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{calmr_basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  collapse = TRUE,
  comment = "#>",
  message = F,
  warning = F
)
```

```{r setup}
library(calmr)
```

# Using calmr in R

## Ingredients

To successfully simulate you only need a data.frame specifying the groups, phases, and trials of the experiment, and (optionally) a data.frame with salience parameters for all the stimuli in the experiment.

## The design data.frame

In the design data.frame, we specify our experimental design. A few rules about the this data.frame:

1. Each row represents a group.
2. Its first column contains the group labels.
3. The remaining columns are organized in pairs. The first column in each pair specifies the trials to give in the phase, whereas the second column in each pair specifies whether the trials within that phase should be randomized.

Let's build a design matrix for a blocking experiment.

```{r}

my_blocking <- data.frame(Group = c("Exp", "Control"),
                          Phase1 = c("10A>(US)", "10C>(US)"),
                          R1 = c(FALSE, FALSE),
                          Phase2 = c("10AB>(US)", "10AB>(US)"),
                          R2 = c(FALSE, FALSE),
                          Test = c("1A#/1B#", "1A#/1B#"),
                          R3 = c(FALSE, FALSE))
my_blocking
```

A few observations about the trial notation:

1. Trials are preceded by a number. That number represents the number of times that trial will be given in each phase. "10A>(US)" means that the "A>(US)" trial will be given 10 times.
2. The presence and absence of the unconditioned stimulus is not denoted with the traditional "+" and "-" symbols. Instead, here we use parenthesis to denote "complex" stimuli. These can be thought of as an element with a complex name (i.e., with more than one character). As such, "(US)" specifies a single element to represent the US.
3. In the same vein, multiple characters with no parentheses denote individual elements. For example, "AB" implies the presence of two stimuli, A and B.
4. The ">" character is used as a separator of the "expectations" and "correction" steps within the trial. "10A>(US)" means that the model generates an expectation with A only, but learns from the co-occurence of A and the US.
5. The "/" character is used as a trial separator (it does not imply randomization by itself). Thus, "1A/1B" specifies that a single "A" trial and a single "B" trial will be given during that phase. Recall that randomization of trials within a phase is specified by the column after it (above, R1, R2, and R3).
6. The "#" character is used to denote probe trials. In contrast to real life, probe trials here entail no update of the model's associations. As such, probe trials can be used to track the development of key associations, with no repercussion to what the model learns on normal training trials.

If you want to check if your trials string will work with the simulator, you can use `trial_parser`. Check `??trial_parser` for more information. The function returns a list with a lot of information, so let's print only some of the fields.

```{r}
trial <- trial_parser("AB/10AC")  #not specifying a number of AB trials. Bad practice!
trial[c("trial_names", "trial_repeats")]
trial <- trial_parser("10AB(AB)(US)") #considering a configural cue for elements AB
trial[c("unique_nominal_stimuli")]

```


## The parameters data.frame

You can run a simulation with only a design data.frame (the simulation function will use default salience values for the stimuli). However, if you want specific salience values for the stimuli in your experiment, you should pass a data.frame with parameters. Let's specify those for our experiment.

```{r}
my_pars <- get_params(my_blocking, model = "HD2022") #Note we can quickly get a table with parameters directly from the design data.frame; a good way to check we are going in the right direction
my_pars$alphas[my_pars$stimulus == "US"] <- .4
my_pars
```

### Additional options

The final ingredient is a list with simulation options. We can get a default one with `get_model_opts`, and modify them accordingly.

```{r}
my_opts <- get_model_opts()
my_opts$iterations <- 3
my_opts

```

### Simulating

And we can finally simulate the model with `quick_model`. By default, the function returns a list with aggregated data. Let's see.

```{r}
my_sim <- quick_model(design_df = my_blocking,
                    param_df = my_pars,
                    options = my_opts,
                    model = "HD2022")
my_sim

```

## Plotting

But, we rarely want the raw data. If you are like me, you will want to see some plots. Let's use `plot` function with the different plots supported by the package (and the model)

Here they are, paired side by side:

### stimulus-specific responding (simple)

The plots below show responding across phases (left to right panels, and trials within). Note, these show the sum of R values (i.e., overall level of responding) per stimulus (absent or present). They do not identify the sources of that responding (i.e., whether A or B is producing the responding); for that, we must look at complex plots.

```{r}
calmr_plot(my_sim, type = "rs", simple = T)
```

### stimulus-specific responding (complex)

These are those complex plots. Here, the parsing changes a little bit, in order to allow comparison among the sources of responding. Now, each stimulus source is shown as a different color, and the vertical panels show different target stimuli.

```{r}
calmr_plot(my_sim, type = "rs", simple = F)
```


### stimulus activations (bar)

stimulus activation plots take us a little deeper into the model. In these, the overall stimulus activation that we use to distribute responding is split into comb and chain activations. Comb activations represent activations due to reciprocal associations (e.g., the effect of US->A on A). Chain activations represent activations via indirect associative links.

Below, each bar is coloured based on the stimulus that generates the activation. The vertical panels denote the target stimulus. The x-axis now shows the trial types.

```{r}
calmr_plot(my_sim, type = "acts", bar = T)
```

### stimulus activations (curve)

Similar to above, but now the x-axis shows trial number, and the trial_types are embedded across horizontal panels.


```{r}
calmr_plot(my_sim, type = "acts", bar = F)
```

### stimulus associations

Going even deeper now. These are the stimulus-stimulus associations that HeiDI uses to generate activations in the first place. Each vertical panel denotes a stimulus source. THe different colours, the target of the association.

```{r}
calmr_plot(my_sim, type = "vs")
```

### stimulus salience at expression

Venture here only if you're ready. The plots below show the stimulus saliencies at the moment of expression.

```{r}
calmr_plot(my_sim, type = "as")
```

## Graphing

You can also take a look at the state of the model's associations at any point during training, using the function `make_graphs`.

```{r}
my_graph_opts <- get_graph_opts("small")
calmr_graph(my_sim, t = 10, graph_opts = my_graph_opts)
```

## Final thoughts

The `calmr` package was designed to simulate quickly; write your design and get a glance of the model predictions. However, the package also has some features for more advanced users, so make sure to check more advanced vignettes when you are ready.
