---
title: "heidi_basics"
author: "Victor Navarro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{heidi_basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  collapse = TRUE,
  comment = "#>",
  message = F,
  warning = F
)
```

```{r setup}
library(heidi)
```

# Using heidi in R

## Ingredients

To successfully simulate you only need a data.frame specifying the groups, phases, and trials of the experiment, and (optionally) a data.frame with salience parameters for all the stimuli in the experiment.

## The design data.frame

In the so-called design data.frame the user specifies their experimental design. A few rules about the this data.frame:

0. Each row represents a group.
1. Its first column contains the group labels.
2. The remaining columns are organized in pairs. The first column in each pair specifies the trials to give in the phase, whereas the second column in each pair specifies whether the trials within that phase should be randomized.

Let's build a design matrix for a blocking experiment.

```{r}

my_blocking <- data.frame(Group = c("Exp", "Control"),
                          Phase1 = c("10A(US)", "10C(US)"),
                          R1 = c(FALSE, FALSE),
                          Phase2 = c("10AB(US)", "10AB(US)"),
                          R2 = c(FALSE, FALSE),
                          Test = c("1A/1B", "1A/1B"),
                          R3 = c(FALSE, FALSE))
my_blocking
```

A few observations about the trial notation:
0. Trials are preceded by a number. That number represents the number of times that trial will be given in each phase. "10A(US)" means that the "A(US)" trial will be given 10 times.
1. The presence and absence of the unconditioned stimulus is not denoted with the traditional "+" and "-" symbols. Instead, we use parenthesis to denote "complex" stimuli. These can be thought of as an element with a complex name (i.e., with more than one character). As such, "(US)" specifies a single element to represent the US.
2. The "/" character is used as a trial separator (it does not imply randomization by itself). Thus, "1A/1B" specifies that a single "A" trial and a single "B" trial will be given during that phase. Recall that randomization of trials within a phase is specified by the column after it (above, R1, R2, and R3).

If you ever want to check if your trials string will work with the simulator, you can use `heidi::trial_parser`. The function returns a list with a lot of information. Check `??trial_parser` for more information. Let's have a go.

```{r}
trial <- trial_parser("AB/10AC")  #not specifying a number of AB trials. Bad practice!
trial[c("trial_names", "trial_repeats")]
trial <- trial_parser("10AB(AB)(US)") #considering a configural cue for elements AB
trial[c("unique_nominal_stimuli")]

```


## The parameters data.frame

You can run a simulation with a design data.frame only (the simulation function will use default salience values for the stimuli). However, if you want specific salience values for the stimuli in your experiment, you should pass a data.frame with parameters. Let's specify those for our experiment.

```{r}
my_pars <- get_params(my_blocking, default_par = .2) #Note we can quickly get a table with parameters directly from the design data.frame; a good way to check we are going in the right direction
my_pars$Alpha[my_pars$Stimulus == "US"] <- .4
my_pars
```

### Additional options

The final ingredient is a list with simulation options. We can get a default one with `get_heidi_opts`, and modify them accordingly.

```{r}
my_opts <- get_heidi_opts()
my_opts$iterations <- 10
my_opts

```

### Simulating

And we can finally simulate the model with `quick_heidi`. By default, the function returns a list with aggregated data. Let's see.

```{r}
my_sim <- quick_heidi(design_df = my_blocking,
                    param_df = my_pars,
                    options = my_opts)
names(my_sim) #contains information about weights (ws), v-values (vs), r-values (rs), and alphas (as)
head(my_sim$vs)

```

## Plotting

But, we rarely want the raw data. If you are like me, you will want to see some plots. Let's use `make_plots` to create some plots, and `patch_plots` to display them.

```{r}
my_plots = make_plots(my_sim)
plot_names = names(my_plots)
plot_names
```

Here they are, paired side by side:

### Stimulus-specific responding (simple)
```{r}
patch_plots(plots = my_plots, selection = plot_names[c(1, 7)]) #Responding
```

### Stimulus-specific responding (complex)
```{r}
patch_plots(plots = my_plots, selection = plot_names[c(2, 8)]) #Responding
```


### Stimulus associations
```{r}
patch_plots(plots = my_plots, selection = plot_names[c(3, 9)])
```

### Combined associative strenghts
```{r}
patch_plots(plots = my_plots, selection = plot_names[c(4, 10)])
```

### Combined associative strenghts (bar)
```{r}
patch_plots(plots = my_plots, selection = plot_names[c(5, 11)])
```

### Stimulus salience at expression
```{r}
patch_plots(plots = my_plots, selection = plot_names[c(6, 12)])
```

## Graphing

You can also take a look at the state of the model's associations at any point during training, using the function `make_graphs`.

```{r}
my_graph_opts <- get_graph_opts("small")
make_graphs(my_sim, trial = 10, opts = my_graph_opts)
make_graphs(my_sim, trial = 20, opts = my_graph_opts)
```

## Final thoughts

The `heidi` package was designed to simulate quickly; write your design and get a glance of the model predictions. However, the package also has some features for more advanced users, so make sure to check more advanced vignettes when you are ready.
